---
title: "Survival_lethal_test"
author: "Sam Gurr"
date: "9/16/2025"
output:
  pdf_document: default
  html_document: default
---

### SET UP
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# SET WORKING DIRECTORY 
knitr::opts_knit$set(root.dir = "C:/Users/gurrs/Documents/Github_repositories/Cgigas_temperature_prime/RAnalysis") # Sam's work
```

### LOAD PACKAGES
```{r, echo=FALSE}
library(ggplot2)
library(tidyr)
library(dplyr)
library(rcompanion)
library(FSA)
library(car)
library(forcats)
library(kableExtra) # nice Rmd tables
library(emmeans)
library(ggpubr)
library(survival)
library(Rmisc)
library(coxme)
library(survminer)
library(ggsurvfit) # survfit2
library(gtsummary) # tbl_survfit
```

# Lethal Test

* Literature review suggests that 44C elicits mortality in C gigas, but what duration?

* Objective: we want to leverage a temperature threshold that elicits mortality, but not all at once. 
Ultimately, our target dataset should yield differences in mortality through time between treatments with different envrionmental histories

* In this 'Lethal test' we exposed 4 half-sib families (N= 5 oysters per family, 20 oysters total per tank) to 30 mins, 1 hour, 2 hours, and 3 hours to 44C
exposed acutely from 20C acclimation

### Test data, load data

```{r load test data, echo=FALSE}

# note: we have percent_survival_rel_embryo and Percent_Survival_rel_Dstage 
# experiment #4 does not have percent relative to embryo, but is percent rel to d hinge stage so we will use that 
# the survival at Dstage as abnormality data can be used for the inital assessment 

# load the raw data, just the counts of dead individuals each day
df_raw       <- read.csv("Data/Lethal_temperature_test/survival_test.csv", header = T) 
df_raw[df_raw == "all"] <- 0
df_raw <- df_raw %>% mutate_at(2:ncol(df_raw), ~as.numeric(as.character(.)))

```

### Test data, for loop for binary format

```{r test data for loop binary format, echo=FALSE}

# loop matrix, start this
loop_columns          <- data.frame(matrix(nrow = 1, ncol = 7))
colnames(loop_columns) <- c('Date', 'Tank_number', 'Treatment', 
                            'Count_Alive', 'Count_Dead', 'Daily_Count_Dead', 'Percent_Survival') # names for comuns in the for loop

# output file from the for loops
df_cols <- data.frame()
df_out  <- data.frame() # output datafarme, start it here

# start the loop
for(r in 1:nrow(df_raw))  { # for each row
  
  loop.date <- df_raw[r,1] # get the date 

  for (c in 2:ncol(df_raw)) { # at each date... call tank columns one by one
    
      # df_raw[c] <- as.numeric(df_raw[c]) # convert to numeric
      
      cumulative.sum <- cumsum(df_raw[c]) # culative sum of the entire column c
      
      cum.count.dead   <- cumulative.sum[r,] # the cumulative count dead at loop.date row r, column c
      
      daily.count.dead <- df_raw[r,c]
      
      count.alive      <- 20 - cum.count.dead # all tanks started with 20 individuals!
      
      percent.survival <- (count.alive/20)*100
      
      tank.num         <- as.numeric(gsub('.*_', '', colnames(df_raw[c])))
      
      
  
  loop_columns$Date         <- loop.date
  
  loop_columns$Tank_number  <- tank.num
  
   if(tank.num == 1 | tank.num == 2 | tank.num == 3) {
    loop_columns$Treatment = "60 minutes"
   } else if (tank.num == 4 | tank.num == 5 | tank.num == 6) {
     loop_columns$Treatment = "120 minutes"
   } else if (tank.num == 7 | tank.num == 8 | tank.num == 9 | tank.num == 10) {
     loop_columns$Treatment = "180 minutes"
   } else {
     loop_columns$Treatment = "30 minutes"
   }
  
  loop_columns$Count_Alive           <- count.alive
  
  loop_columns$Count_Dead            <- cum.count.dead # cumulative sum of count dead
  
  loop_columns$Daily_Count_Dead      <- daily.count.dead # cumulative sum of count dead
  
  loop_columns$Percent_Survival      <- percent.survival 
  
  df_cols      <- rbind(df_cols,loop_columns) #bind to a cumulative list dataframe
  
  # Note: df_cols now contains as many rows as there are tank IDs at the given date loop.date
  
  } # close columns for each tank
    df_out      <- rbind(df_out,df_cols) 
    df_cols      <- data.frame() # start columns over again, gets the next row (date) for te previous loop
    print(df_out) # print to monitor progress
    
} # close rows for each date - output cumulative table row by row for each tank
View(df_out)
```

### Test data, plot as barplots mean +- SE for each run
 
```{r test data summary stats}

SummaryStats <- summarySE(df_out, measurevar="Percent_Survival", groupvars=c("Treatment", "Date"))


# plot relative to stocking density aas embryos 



# plot relative to 100$% restart at D stage 
PlotMeanSE <-  ggplot(SummaryStats, aes(x=Date, y=Percent_Survival, color=Treatment)) + geom_line()+
                          geom_point()+ 
                           geom_errorbar(aes(ymin=Percent_Survival-se, 
                                             ymax=Percent_Survival+se), width=.2,
                                        position=position_dodge(.1)) +
                          # scale_color_manual(values=c("green4", "darkorange1", "purple"))+
                          theme(legend.position="right", legend.direction="vertical", 
                                legend.title = element_blank())+ theme_bw() + 
                          labs(x="Date", y="Percent Survival")+ #scale_y_continuous(breaks=seq(0,100,20))+ 
                          ggtitle("Survival following 44C exposure (under 20C)") + 
                          theme(panel.grid.major = element_blank(), 
                                panel.grid.minor = element_blank(), 
                                panel.background = element_blank(), 
                                axis.line = element_line(colour = "black"))
PlotMeanSE


# output the data !!

write.csv(SummaryStats,"Output/survival_test_MeanSE.csv")

pdf("Output/survival_test_MeanSE.pdf", width=10, height=5)
print(PlotMeanSE)
dev.off()

```

### Test data, Kaplan Meier

#### About:

* We will use the {ggsurvfit} package to generate Kaplan-Meier plots. This package aims to ease plotting of time-to-event endpoints using the power of the {ggplot2} package. See http://www.danieldsjoberg.com/ggsurvfit/index.html for details.

#### Build 'surv' object for statistical analysis 

* useful links [https://www.emilyzabor.com/tutorials/survival_analysis_in_r_tutorial.html#Comparing_survival_times_between_groups]

* What the hell does a surv object mean?? 
  
  - a 'surv' object reads ALL events as that of a dead (1) or alive (0) individual
  
  - but but but... we have volumetric count data of live and dead! Does this not throw a wrench into this type of analysis? 
  
    - Nope! Below we use logic to create a GIGANTIC table formatted correctly for this purpose!

#### Format

 * Build new column for count dead - take the count alive data from time 0 and porp survival to calcualted number dead

#### Convert to binary formant (super long data tbale!) using two for loops:

 * (1) Including tank - to use as surv object in coxme statistics (random factor included!) not suitable for plotting
 
 * (2) Mean counts alive and dead by tank - better represented for plots!


```{r test data KM data}
# create new columns for live and dead based on the count and proportion (of alive) tobackcalculate dead - these are integers! 
df_KM <- df_out %>% 
          dplyr::mutate(Count_Alive = 
                          as.numeric(gsub(",","",Count_Alive))) %>% # remove the commas from these numbers- id as character if not removed!
          dplyr::mutate(proportion = Percent_Survival/100,
                        Count_total = case_when(proportion == 0 ~ 0,
                                                proportion >0 ~
                                                 (1/as.numeric(proportion))*
                                                 as.numeric(Count_Alive)))# %>% # proportion is that of alive
                           #Count_dead = Count_total - Count_Alive)
          # dplyr::filter(!proportion == 0.00) # omit rows where the number of live individuals is 0

View(df_KM)
```

### Test data, convert to binary format

```{r test data convert to binary}

# (1) Including tank - to use as surv object in coxme statistics (random factor included!) not suitable for plotting

# Call the cumulative dataframe that we will write to in the for loop below
df_binary     <- data.frame() # start dataframe  - all
# run it 
for (i in 1:nrow(df_KM)) {
   #count_alive <- round(df_Exp2[i,]$Count_alive/100) # divide all counts by 1000, round to nearest
   #count_dead  <- round(df_Exp2[i,]$Count_dead/100)  # divide all counts by 1000, round to nearest
   
   count_alive <- round(df_KM[i,]$Count_Alive) # raw counts
   count_dead  <- round(df_KM[i,]$Daily_Count_Dead)  # raw counts
   
   loopDF           <- data.frame(matrix(0,ncol = 4, nrow = (count_alive + count_dead))) 
   colnames(loopDF) <- (c('Date','Treatment','Tank_number','Count_dead'))
   loopDF           <- loopDF %>% 
                            dplyr::mutate(
                              Treatment   = df_KM[i,]$Treatment,
                              Tank_number = df_KM[i,]$Tank_number,
                              Date        = df_KM[i,]$Date)
  loopDF[c(1:count_dead),4] = 1 # 1 for dead 
  loopDF[c((count_dead + 1):nrow(loopDF)),4] = 0 # 0 for alive

  loopDF           <- data.frame(loopDF) # name dataframe for this single row
  df_binary <- rbind(df_binary,loopDF) #bind to a cumulative list dataframe
}

df_binary <- df_binary %>% na.omit()
# View(df_binary)

# Convert to Date class
dates_asDate <- as.Date(df_binary$Date, format = "%Y%m%d")

# Find earliest date
earliest <- min(dates_asDate)

# Calculate number of days since earliest date
days_since_earliest <- as.numeric(dates_asDate - earliest)

# 
df_binary$Days <- as.numeric(days_since_earliest)
```

### Test data, breat survival object

```{r inital test, create surv object}

surv_all <- survfit2(Surv(Days, Count_dead) ~ Treatment, data = df_binary)

```

### Test data, plot Kapla Meier

```{r test data, plot Kaplan Meier}

KaplanMeier.plot <- surv_all %>% 
                        ggsurvfit(linetype_aes = TRUE, linewidth = 1,type ="survival") +
                        scale_color_manual(values = c("lightgreen","turquoise","orange","darkred")) +
                        scale_fill_manual(values = c("lightgreen","turquoise","orange","darkred")) +
                        labs(
                          x = "Days",
                          y = "Overall survival probability"
                        ) + 
                        add_confidence_interval() +
                        # add_risktable() +
                        # add_risktable_strata_symbol() + # (symbol = "\U25CF", size = 10
                        scale_ggsurvfit(x_scales = list(breaks = 0:10)) #+
                       # add_pvalue(location  = "annotation", x = 8.5)

# plot the surv data
pdf("Output/survival_test_KaplanMeier.pdf", width=5, height = 4)
KaplanMeier.plot
dev.off()

```
